---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fishbc

<!-- badges: start -->
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![R build status](https://github.com/poissonconsulting/fishbc/workflows/R-CMD-check/badge.svg)](https://github.com/poissonconsulting/fishbc/actions)
[![License: CC0](https://img.shields.io/badge/License-CC0-green.svg)](https://tldrlegal.com/license/creative-commons-cc0-1.0-universal)
[![CRAN status](https://www.r-pkg.org/badges/version/fishbc)](https://CRAN.R-project.org/package=fishbc)
<!-- badges: end -->

## Introduction

`fishbc` is an R package that provides provides raw and curated data on the codes, classification and conservation status of freshwater fishes in British Columbia. 
Marine fishes will be added in a future release.

It contains information licensed under the [Open Government Licence – British Columbia](https://www2.gov.bc.ca/gov/content/data/open-data/open-government-licence-bc).

## Installation

To install the latest development version from [GitHub](https://github.com/poissonconsulting/fishbc)
```r
# install.packages("remotes")
remotes::install_github("poissonconsulting/fishbc")
```

## Demonstration

```{r example}
library(tibble) # for tidy printing
fishbc::freshwaterfish
```

## Contribution

Please report any [issues](https://github.com/poissonconsulting/fishbc/issues).

[Pull requests](https://github.com/poissonconsulting/fishbc/pulls) are always welcome.

## Code of Conduct

Please note that the fishbc project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). 
By contributing to this project, you agree to abide by its terms.

