---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fishbc

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![R build status](https://github.com/poissonconsulting/fishbc/workflows/R-CMD-check/badge.svg)](https://github.com/poissonconsulting/fishbc/actions)
[![codecov](https://codecov.io/gh/poissonconsulting/fishbc/branch/master/graph/badge.svg?token=2EvrLVvDcZ)](https://codecov.io/gh/poissonconsulting/fishbc)
[![License: CCBY4](https://img.shields.io/badge/License-CCBY4-green.svg)](https://creativecommons.org/licenses/by/4.0/)
[![CRAN status](https://www.r-pkg.org/badges/version/fishbc)](https://CRAN.R-project.org/package=fishbc)
![CRAN downloads](https://cranlogs.r-pkg.org/badges/fishbc)

[![R-CMD-check](https://github.com/poissonconsulting/fishbc/workflows/R-CMD-check/badge.svg)](https://github.com/poissonconsulting/fishbc/actions)
<!-- badges: end -->

## Introduction

`fishbc` is an R package that provides provides raw and curated data on the codes, classification and conservation status of freshwater fishes in British Columbia. 
Marine fishes will be added in a future release.
It also include raw fish codes for BC Conservation Data Centre and the Government of Alberta.

It contains information licensed under the [Open Government Licence – British Columbia](https://www2.gov.bc.ca/gov/content/data/open-data/open-government-licence-bc) and the [Open Government Licence - Alberta](https://open.alberta.ca/licence)

## Installation

To install the latest release from [CRAN](https://cran.r-project.org)
```r
install.packages("fishbc")
```

To install the latest development version from [GitHub](https://github.com/poissonconsulting/fishbc)
```r
# install.packages("remotes")
remotes::install_github("poissonconsulting/fishbc")
```

## Demonstration

Get the curated freshwater fish of BC fish codes.

```{r}
library(tibble) # for tidy printing
fishbc::freshwaterfish
```

Get the common names for fish codes.

```{r}
fishbc::fbc_common_name(c("AF", "WSG", NA, "AF", "NOTACODE"))
```

## Contribution

Please report any [issues](https://github.com/poissonconsulting/fishbc/issues).

[Pull requests](https://github.com/poissonconsulting/fishbc/pulls) are always welcome.

## Code of Conduct

Please note that the fishbc project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). 
By contributing to this project, you agree to abide by its terms.

